### 2016-06-06 paper.tex
Initial bits

\documentclass[a4paper]{article}

\usepackage{natbib}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}

\usepackage{tikz}
\usetikzlibrary{math,fpu,calc,fit,mindmap,backgrounds,positioning}

\usepackage{xspace}
\newcommand{\eg}{e.g.\xspace}
\newcommand{\etal}{et al.\xspace}
\newcommand{\ie}{i.e.\xspace}
\newcommand{\etc}{etc.\xspace}
\newcommand{\vs}{\textit{vs.}\xspace}

\title{From Social Attention to Social Modelling in Robotics:\\ The Robotic Social Awareness Model}

\author{Séverin Lemaignan}

\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Abstract goes here
\end{abstract}

\section{What do we aim for?}

Human social dynamics rely upon the ability to effectively attribute beliefs,
goals and percepts to other people. This set of meta-representational abilities
shapes what is called a theory of mind (ToM) or the ability to mentalize, and
leads to mutual modelling: the reciprocal ability to establish a mental model of
the other. This lays at the core of human interactions: normal human social
interactions depend upon the recognition of other sensory perspectives, the
understanding of other mental states, and the recognition of complex non-verbal
cues of attention and emotional state. As such, adapting and transferring these
cognitive skills to social robots is an important research objective.

Understanding and predicting one's behaviour.

\subsection{Surface Alignment vs Deep Grounding}

\citet{pickering2006alignment} argue that mutual understanding
starts mostly with a \emph{superficial alignment} at the level of the linguistic
representations, due to priming mechanisms, and that this local alignment may --
in some cases -- lead to a \emph{global alignment} of the semantic level
(\emph{deep grounding}).  For these authors, the convergence in dialogue, and
even the repair of some misunderstandings, is explained by this mimetic behavior
more than by a monitoring of each other's knowledge:

\begin{quote}
[\ldots] interlocutors do
not need to monitor and develop full common ground as a regular, constant
part of routine conversation, as it would be unnecessary and far too costly.
Establishment of full common ground is, we argue, a specialized and
non-automatic process that is used primarily in times of difficulty (when
radical misalignment becomes apparent).~\citep[p.179]{pickering2006alignment}
\end{quote}

\subsection{Theory of Mind}

In~\cite{scassellati2002theory}, Scassellati gave
an initial account of Leslie's and Baron-Cohen's respective models of the
emergence of a theory of mind (we discuss them below) from the perspective of
robotics, but reported implementation work was limited to simple perceptual
precursors (like face detection or color saliencies detection). Since then,
research in this field has been focused on applications relying on Flavell's
\emph{Level 1}~\cite{flavell1977development} perspective-taking, \ie
perspective-taking that only requires perceptual abilities (``\emph{I see (you do
not see the book)}''), and actually mostly limited to visual perception (relevant
work include Breazeal~\cite{breazeal2006using}, Trafton~\cite{Trafton2005} and
Ros~\cite{Ros2010}).

Based on perspective taking \emph{Level 1} alone, Breazeal
\etal\cite{breazeal2009embodied} and Warnier \etal\cite{warnier2012when}
successfully tackled the classical hallmark of theory of mind, the false belief
experiment (also known as the ``Sally and Anne'' experiment,
Figure~\ref{false-beliefs}, introduced by~\cite{wimmer1983beliefs}, original
experimental setting by~\cite{baron1985does}). They demonstrated complete
human-robot interaction scenarios where robots recognize and handle false belief
situations in dyadic or triadic interactions, and exhibit helping behaviours
that account for the missing/false beliefs of the human partners.

\section{A Cognitive Model of Social Awareness for Robots}

\subsection{The Attention Schemata Theory}

Our approach draws form the \emph{Attention Schemata Theory}, proposed by
Grazino~\cite{graziano2013consciousness}.

What awareness can do? ``the brain \emph{does} attention, but \emph{knows}
awareness'': as such, ``awareness can in principle be verbally reported''.

\subsection{Associative Memory Network}

In particular, used for behavioural alignment

\subsection{The Robotic Social Awareness Model}

Associative Memory as an Informational Proxy for the Attention System

\begin{enumerate}
    \item Perceptual inputs feed in an associative memory
    \item the resulting set of pre-activated units is labelled as the current
        \emph{focus of attention}
    \item According to the Attention Schemata Theory, by \emph{explicitely}
        labelling these units as being attended, we make the robot \emph{aware}
        of the corresponding phenomenons.
\end{enumerate}

\section{Attention}

Modelled as a \emph{Biased Competition Model of
Attention}~\cite{desimone1995neural}.
Implemented using a particular \emph{Associative Memory Network} with an
additional top-down biasing mechanism.

\begin{figure}
    \centering
        \resizebox{0.7\linewidth}{!}{%
        \begin{tikzpicture}[
                >=latex,
                every edge/.style={draw, very thick},
                every node/.style={font=\bf,align=center},
                input/.style={draw,circle, inner sep=0pt,minimum size=0.6cm}
            ]

            % draw the units
            \foreach \x [count=\xi] in {0, 120, 240}
                \foreach \y [count=\yi] in {-.6, 0, .6}
                    %\node[input, ball color=orange!\x!blue] at ($(\x:3)!\y!90:(0,0)$) (i\xi\yi) {};
                    \node[input, fill=blue!\x!green] at ($(\x:3)!\y!90:(0,0)$) (i\xi\yi) {};

            \node[draw,rounded corners, fit=(i11)(i12)(i13)] (modA) {};
            \node[draw,rounded corners, rotate fit=120,fit=(i21)(i22)(i23)] (modB) {};
            \node[draw,rounded corners, rotate fit=60,fit=(i31)(i32)(i33)] (modC) {};

            \node at (180:5.5) {Modalities} edge[->] (modB) edge[->] (modC);
            \node at (-40:6) {Input\\units} edge[->] (i13) edge[->] (i31);

            \path[draw,dashed, thick]  (i23) -- (i11) -- (i31) -- (i23);
            \path[draw,dotted, thick] (i23) -- (i12) -- (i32) -- (i23);
            \path[draw] (i21) -- (i13) -- (i33) -- (i21);

            \node at (1,1) {Associative\\links};
        \end{tikzpicture}
    }

    \caption{Associative Memory Network}
    \label{blabla}
\end{figure}

\subsection{Biasing Mechanisms}

Biasing competition~\cite{beck2009top}

The bottom-up biasing mechanism follows naturally from the structure of the
associatve memory model: a strong and long activation of a perceptual input
leads to the activation of the corresponding unit in the memory network and the
suppression of competing inputs.

The top-down mechanism is to be invented :-)
It should enable high-level decisional processes to effectively suppress (or
reinforce) units. The \emph{nature} and \emph{representation} of these
high-level processes is unclear, but might be of symbolic nature.

\section{Social Attention}

What do we call the \emph{attention state} of a partner?

Grazino~\cite{graziano2013consciousness}: ``the mental machinery to model
someone else attentional state is the same as the one used for oneself.''
``In both case, the purpose is understanding and predicting one's behaviour.''

Cues from which we reconstruct someone else's attentional state (from Graziano):
\begin{itemize}
    \item gaze direction
    \item facial expression
    \item body language
    \item prior knowledge of person
    \item location of salient objects
\end{itemize}

\section{The functional pre-requisites for social robots}

\subsection{Social Assessment}
\subsubsection{Gaze direction}
\subsubsection{Facial and body expressions}
\subsubsection{Gestures}

\subsection{Situation Assessment}

Including saliency information

\section{Evaluation of the Model}

\subsection{Does it perform as well as Symbolic Architectures?}
\subsection{Comparison with Symbolic Situation Assessment}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}

### 2016-06-06 paper.tex
A lot of meat

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{What do we aim for?}
@@
 
+An integrated theory of social cognition, applicable to robots
+
+A principled yet workable approach to social cognition for artificial intelligence
+
+\subsection{Problem Statement}
+
+Symbolic approaches to social cognition work by first building a mental model of the
+interacting humans. This is typically done by a combination of 3D situation
+assessment (the robot builds and update a semantic 3D model of its environment)
+and visual perspective taking (based on the estimation of the pose of the human
+head). This permits the computation of allocentric, and more importantly,
+egocentric spatial relations between the spatial entities in the environment
+(we call it \emph{social situation assessment}).
+
+This geometric computations are then turned into symbolic representations,
+typically using logical statements (embedded in logic
+programming~\cite{tenorth2009knowrob} or ontologies~\cite{lemaignan2010oro}).
+
+The robot creates and continuously updates one symbolic model per
+agent~\cite{lemaignan2010oro}. These models are then used by other cognitive
+processes (task planning, dialogue, task execution supervision) that are
+designed to take advantage of the agents' knowledge models to produce
+socially-aware behaviours: for example, the task planner may plan manipulation
+task using only entities visible to the human~\cite{lallement2014hatp}, or the
+dialogue manager may use the specific knowledge model of the speaker to
+interpret the speech, avoiding grounding ambiguities that might otherwise
+occur~\cite{lemaignan2011grounding}.
+
+This works well as long as we limit ourselves to the manipulation of the results
+of visual perspective taking. However, one intuitively recognizes that social
+modelling goes indeed beyond computing what the human perceives or does not
+perceive. This has been clearly recognized in developmental psychology, for
+instance with Flavell's distinction between \emph{cognitive connections} on one
+hand, and \emph{mental representations} on the other hand (we will come back to
+Flavell's \emph{connection-representation} account in
+section~\ref{connection-representation}). Now, if we are to model someone else's
+mind beyond a naive geometric model of their perception, we indeed enter the
+realm of \emph{representations}. What are they? How to access them? How to
+represent and manipulate them? These three questions lay at the core of this
+work, and as such we effectively take over the point we previously made
+in~\cite{lemaignan2015mutual}:  we ultimately want to come up with a
+meta-representational cognitive system to \emph{represents
+representations} (including representations of \emph{false} beliefs or unknown
+facts, \ie suppositions).
+
+We must immediately clarify that, even though this goal may seem to pre-suppose
+\emph{symbolic} meta-representations, this is not the case: at that stage, we do
+not have evidence that a particular kind of computational structure may better
+support the representation and manipulation of mental representations.
+
+\subsection{From Social Attention to Social Modelling}
+
+The hypothesis that we hereafter develop and turn into a cognitive model is the
+following: mental representations are snapshots of awareness, awareness being
+itself a label for the memory-mediated process of attention.
+
 \subsection{Surface Alignment vs Deep Grounding}
 
-\citet{pickering2006alignment} argue that mutual understanding
+Pickering and Garrod~\cite{pickering2006alignment} argue that mutual understanding
 starts mostly with a \emph{superficial alignment} at the level of the linguistic
@@
 
+We previously covered the state of the research on theory of mind in social
+robotics~\cite{lemaignan2015mutual}.
+
 In~\cite{scassellati2002theory}, Scassellati gave
@@
 successfully tackled the classical hallmark of theory of mind, the false belief
-experiment (also known as the ``Sally and Anne'' experiment,
-Figure~\ref{false-beliefs}, introduced by~\cite{wimmer1983beliefs}, original
-experimental setting by~\cite{baron1985does}). They demonstrated complete
-human-robot interaction scenarios where robots recognize and handle false belief
-situations in dyadic or triadic interactions, and exhibit helping behaviours
-that account for the missing/false beliefs of the human partners.
+experiment (also known as the ``Sally and Anne'' experiment, introduced
+by~\cite{wimmer1983beliefs}, original experimental setting
+by~\cite{baron1985does}). They demonstrated complete human-robot interaction
+scenarios where robots recognize and handle false belief situations in dyadic or
+triadic interactions, and exhibit helping behaviours that account for the
+missing/false beliefs of the human partners.
 
-\section{A Cognitive Model of Social Awareness for Robots}
+\cite{devin2016implemented}
 
-\subsection{The Attention Schemata Theory}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Foundations of the Socio-cognitive Awareness for Robots}
 
-Our approach draws form the \emph{Attention Schemata Theory}, proposed by
+\subsection{Connections vs. representations}
+\label{connection-representation}
+
+In~\cite{flavell1990developmental}, Flavell relates perspective taking
+\emph{Level 1} to establishing \emph{cognitive connections} (I see, I hear, I
+want, I like, I fear...), in contrast to perspective taking \emph{Level 2} that
+relates to manipulating \emph{representations}.  This is exemplified by
+\emph{appearance-reality} tasks, like the \emph{elephant mask} experiment
+proposed in~\cite{flavell1990developmental}: 3-years old children are not able
+to tell that an experimenter hidden behind a large elephant mask but who speaks
+normally \emph{looks} like an elephant, \emph{sounds} like the experimenter, and
+\emph{really is} the experimenter.  It appears that, while those children are
+able to explicitly manipulate cognitive connections (they know for instance that
+these are largely independent of each other and that they can evolve over time)
+and know as well that their own connections are independent of those of other
+people, they do not think that one concept can \emph{seriously} (\ie non
+playfully) hold several, possibly conflicting, representations.
+
+\subsection{Phenomenal \vs Access Consciousness}
+
+Neuroscientist view: Block's proposal of two kind of consciousness~\cite{block1996can}:
+\emph{phenomenal} consciousness as the raw inner experience; \emph{access}
+consciousness as the more abstract, cognitive ability to think about and report
+on those experiences.
+
+Does this map to the traditional sub-symbolic/symbolic divide that we observe in
+artificial intelligence, and in particular in robotics?
+
+The \emph{phenomenal consciousness} would then be the immediate raw perceptual
+inputs: video streams from cameras, readings for torque and force sensors, joint
+positions, etc.
+
+The \emph{access consciousness} would typically be the symbolic representation
+of these inputs.
+
+We must however keep in mind that there is likely no such rigid dichotomy
+between phenomenal consciousness and access consciousness. It is rather a
+continuum of processing~\citep[p.55]{graziano2013consciousness}
+
+Recursivity of consciousness: if I'm looking at a green apple, my cognitive
+machinery can decide and report that I'm aware of green. I can also be aware of
+the deciding and aware of the reporting~\citep[p.55]{graziano2013consciousness}.
+
+\subsection{The Attention Schema Theory}
+
+Our approach draws form the \emph{Attention Schema Theory}, proposed by
 Grazino~\cite{graziano2013consciousness}.
@@
 
-\subsection{Associative Memory Network}
+\subsection{Associative Memory Networks}
 
@@
 
-\subsection{The Robotic Social Awareness Model}
+\subsection{Semantic Networks}
+
+Ontologies and other Knowledge Representation and Reasoning systems.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{The Robotic Social Awareness Model}
+
+\subsection{Overview}
 
@@
     \item Perceptual inputs feed in an associative memory
-    \item the resulting set of pre-activated units is labelled as the current
-        \emph{focus of attention}
-    \item According to the Attention Schemata Theory, by \emph{explicitely}
-        labelling these units as being attended, we make the robot \emph{aware}
-        of the corresponding phenomenons.
+
+    \item the resulting set of pre-activated units is the current
+        \emph{focus of attention} -- this step may require dimensionality
+        reduction techniques
+
+    \item According to the Attention Schema Theory, by \emph{explicitely}
+        labelling the activated and pre-activated units as being attended, we
+        make the robot \emph{aware} of the corresponding phenomenons.
+
+    \item The labelling occurs by the mean of a symbolic association within a
+        semantic network (ontology)
+
 \end{enumerate}
 
-\section{Attention}
+\subsection{Attention}
 
@@
 
-\subsection{Biasing Mechanisms}
+\subsubsection{Biasing Mechanisms}
 
@@
 
-\section{Social Attention}
+\subsection{Social Attention}
 
@@
 
-\section{The functional pre-requisites for social robots}
+\subsection{What is the contribution, the novelty?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{What Does the Model Predict?}
+
+\subsection{Behavioural Alignment}
+\subsection{Verbalization of the (Social) Attentional Process}
+\subsection{False-beliefs}
+
+\subsection{Symbol Grounding}
+
+\subsection{Recursive Awareness}
+
+I can describe what I'm aware of, I can also recursively be aware I'm describing
+my state of awareness, and I can verbalize this second order awareness.
+
+This should work \emph{Attention Schema} theory because the attention schema
+is seen as a regular sensory
+representation~\citep[p.55]{graziano2013consciousness}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{The Functional Pre-Requisites for Social Robots}
 
 \subsection{Social Assessment}
-\subsubsection{Gaze direction}
-\subsubsection{Facial and body expressions}
+\subsubsection{Gaze Direction}
+\subsubsection{Facial and Body Expressions}
 \subsubsection{Gestures}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Evaluation of the Model}

### 2016-07-01 paper.tex
The social hypothesis
WIP

@@
 \begin{abstract}
-Abstract goes here
+
+My current perspective could be summarized as:
+
+1- mental representations are snapshots of what we are aware of;
+
+2- awareness is the label we conveniently put on the process of attention;
+
+3- attention at time t is the label we put on the set of the activated units in
+a (biased) associative memory network;
+
+4- modelling others’ mental representations is taking snapshots of their own
+current state of awareness;
+
+5- modelling other’s state of awareness, ie, their current attentional process,
+is mediated by one’s own attentional system, typically through joint attention
+mechanisms;
+
+6- Points 1 to 5 essentially refer to a 'phenomenal' awareness (a 'raw' inner
+experience). 'Phenomenal' awareness can be turned into 'access' consciousness
+(the abstract, cognitive ability to reflect on the inner experience);
+
+7- In robots, 'access consciousness' can be mapped to symbolic, logic
+representations.
+
+(2) is Graziano's Attention Schemata Theory; (3) is based on Biased Competition
+Model of Attention, Desimone and Duncan 1995; (6) comes from Block 1996.
+
+As you hopefully see, this drafts a sub-symbolic/symbolic hybrid model of
+cognition that puts (social) attention at the centre of social cognition.
+
+Some of the main 'gaps' are:
+
+- mental representations are likely not simply instantaneous 'snapshots' --
+even though I could argue that they indirectly embed a notion of time and
+history as they are snapshots of activated units in an associative memory
+network;
+
+- equating attention to a set of activated units in a memory network raises
+questions regarding the nature of these units: physical entities? active
+interaction modalities? ...? In particular, the right level of abstraction of
+these units is not immediately clear: the spectrum is rather large, from raw
+perceptual inputs 'à la Tony Morse' to high-level units like objects, joint
+gazing, etc.
+
+- the Biased Competition Model of Attention supports interesting bottom-up and
+top-down biasing mechanisms: bottom-up is quite clear (if a unit is activated
+longer/stronger, it biases the resulting attention to this unit. Top-down is
+less clear (but also potentially very interesting as it closes the loop between
+the sub-symbolic and symbolic models): how more abstract cognitive process can
+influence the memory network to bias the attention process?
+
+- the access to other's mental representations is mediated by one's own
+attentional system: the consequences of this mediation are not immediately
+clear
+
+- I conveniently map the split between 'phenomenal consciousness' and 'access
+consciousness' respectively to sub-symbolic and symbolic computational
+structures. This need to be better evidenced, if only because the frontier
+between phenomenal and access consciousness is all be clear (as pointed by
+Graziano).
+
+- It is not yet clear how the sub-symbolic and symbolic models of the different
+agent co-exist: is each agent a different unit in the same associative network?
+one associative memory per agent? what about the symbolic level?
+
+Now, as we discussed many times, the model is of little interest if it does not
+have some predictive power.
+
+Here a few social behaviours that I would like a robot to exhibit if endowed
+with such a model -- Not proper predictions, though.
+
+- behavioural alignment -> surface alignment and/or global alignment -> the
+original 'maze' task by Pickering and Garrod is certainly relevant here: every
+time you play the same game with the same person, the interaction is more fluid
+- ability to pass false-belief tasks, including non-physical, abstract ones, -
+recursive awareness: being aware of being aware -- typically evidenced by being
+able to describe/verbalize its own state of awareness - 'natural' turn-taking:
+I 'just' know when it is my turn to act - 'natural' protodeclarative pointing:
+I 'just' know when I really need to draw your attention on something
+
+The word 'just' is clearly the key here: these capabilities should be natural
+in the sense that they naturally follow from the model.
+
 \end{abstract}
@@
 
-A principled yet workable approach to social cognition for artificial intelligence
+A principled yet practical approach to social cognition for artificial intelligence
 
@@
 
+\subsection{What is not adequately solved by current techniques?}
+
+[give concrete examples of social situations that are not easily achieveable
+with current (symbolic or not) approaches]
+
+[this examples should be turned into predictions for what our proposal should be
+able to achieve]
+
 \subsection{From Social Attention to Social Modelling}
@@
 The hypothesis that we hereafter develop and turn into a cognitive model is the
-following: mental representations are snapshots of awareness, awareness being
-itself a label for the memory-mediated process of attention.
+following: \textbf{mental representations are snapshots of awareness, awareness being
+itself a label for the memory-mediated process of attention}.
+
+This extends to social cognition: \textbf{modelling others' mental representations is
+taking snapshots of their current state of awareness}. As we do not have direct
+access to others' process of attention, it has to be mediated. We suggest that
+\textbf{modelling other's state of awareness is mediated by one's own
+attentional system, through joint attention mechanisms}.
+
+This results in the following set of hypotheses:
+
+\begin{itemize}
+    \item \textbf{H1}: 
+\end{itemize}
 
@@
 \end{quote}
+
+analysis of behavioural alignment between partners (via
+metrics like the recently proposed \emph{Individual Motor
+Signature}~\cite{slowinski2016dynamic})
 
@@
 
-\begin{figure}
+\subsubsection{Memory-mediated Attention}
+
+\emph{Attention} is modelled as the set of activated units in an associative
+memory network.
+
+Interplay between multi-modal social cues on one hand, and priming of previously activated physical
+entities (objects, agents) and
+
+Attention
+
+\begin{figure}[h!]
     \centering
@@
 someone else attentional state is the same as the one used for oneself.''
+
 ``In both case, the purpose is understanding and predicting one's behaviour.''
@@
 
+\subsection{Protodeclarative pointing}
+
+Protoimperative vs protodeclarative pointing~\cite{baron1986perceptual}
+
 \subsection{Behavioural Alignment}
@@
 \subsection{False-beliefs}
+
+Picture ordering protocol, with \emph{mechanical}, \emph{behavioural} and
+\emph{intentional} sequences~\cite{baron1986mechanical}
 
@@
 
+\subsection{Adapted behaviour to the interacting agent}
+
+A robot facing a baby, a 3 years old, a 5 years old, a 13 years old or an adult
+should behave differently.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Social Assessment}
+
 \subsubsection{Gaze Direction}

### 2017-03-27 paper.tex
Re-structuring the notes on the model of social awareness in a paper on freeplay

--- 2016-06-06 paper.tex
@@
 
-\title{From Social Attention to Social Modelling in Robotics:\\ The Robotic Social Awareness Model}
+\graphicspath{{figs/}}
 
-\author{Séverin Lemaignan}
+\title{The robot who pretend-played with me}
+
+\author{Séverin Lemaignan, Tony Belpaeme}
 
@@
 
+Alternative titles:
+
+\begin{itemize}
+    \item Can I have real fun playing with my robot?
+    \item A Window Onto Your Mind: Playful Robots Leveraging Attention to
+        Predict Actions
+    \item Playing Like your Friend: Dealing with the High Dimensionality of
+        Real-World Social Dynamic
+
+\end{itemize}
+
 \begin{abstract}
-Abstract goes here
+
+    This article presents how attention tracking enables a robot to first automatically segment
+    a social interaction into meaningful situations (\emph{situation awareness}), and
+    second, to predict the participants' actions.
+
+    We study this effect in a high-dimensional social interaction situation based on
+    free dramatic play: pairs of children (4-7 years old) or robot and child are
+    invited to freely interact with characters and items on an interactive
+    table, telling and acting stories with no explicit preset goal.
+
+    Although the actions performed by the children are not known beforehand, our
+    approach allows the robot to segment on the fly the interaction flow into
+    episodes whose main focus can be predicted and used to generate pro-social
+    behaviours.
+
 \end{abstract}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{What do we aim for?}
+\section{Introduction}
+
+\subsection{Mutual Modelling: Theoretical Background}
 
@@
 
-A principled yet workable approach to social cognition for artificial intelligence
+A principled yet practical approach to social cognition for artificial intelligence
 
-\subsection{Problem Statement}
+\subsubsection{Pretense}
+
+A second direction for the 'meta-cognitive foundations of mind modelling' is
+Leslie's paper on the role of pretense~\cite{leslie1987pretense}.
+
+\subsubsection{Surface Alignment vs Deep Grounding}
+
+Pickering and Garrod~\cite{pickering2006alignment} argue that mutual understanding
+starts mostly with a \emph{superficial alignment} at the level of the linguistic
+representations, due to priming mechanisms, and that this local alignment may --
+in some cases -- lead to a \emph{global alignment} of the semantic level
+(\emph{deep grounding}).  For these authors, the convergence in dialogue, and
+even the repair of some misunderstandings, is explained by this mimetic behavior
+more than by a monitoring of each other's knowledge:
+
+\begin{quote}
+[\ldots] interlocutors do
+not need to monitor and develop full common ground as a regular, constant
+part of routine conversation, as it would be unnecessary and far too costly.
+Establishment of full common ground is, we argue, a specialized and
+non-automatic process that is used primarily in times of difficulty (when
+radical misalignment becomes apparent).~\citep[p.179]{pickering2006alignment}
+\end{quote}
+
+analysis of behavioural alignment between partners (via
+metrics like the recently proposed \emph{Individual Motor
+Signature}~\cite{slowinski2016dynamic})
+
+\subsubsection{Theory of Mind}
+
+We previously covered the state of the research on theory of mind in social
+robotics~\cite{lemaignan2015mutual}.
+
+In~\cite{scassellati2002theory}, Scassellati gave
+an initial account of Leslie's and Baron-Cohen's respective models of the
+emergence of a theory of mind (we discuss them below) from the perspective of
+robotics, but reported implementation work was limited to simple perceptual
+precursors (like face detection or color saliencies detection). Since then,
+research in this field has been focused on applications relying on Flavell's
+\emph{Level 1}~\cite{flavell1977development} perspective-taking, \ie
+perspective-taking that only requires perceptual abilities (``\emph{I see (you do
+not see the book)}''), and actually mostly limited to visual perception (relevant
+work include Breazeal~\cite{breazeal2006using}, Trafton~\cite{Trafton2005} and
+Ros~\cite{Ros2010}).
+
+Based on perspective taking \emph{Level 1} alone, Breazeal
+\etal\cite{breazeal2009embodied} and Warnier \etal\cite{warnier2012when}
+successfully tackled the classical hallmark of theory of mind, the false belief
+experiment (also known as the ``Sally and Anne'' experiment, introduced
+by~\cite{wimmer1983beliefs}, original experimental setting
+by~\cite{baron1985does}). They demonstrated complete human-robot interaction
+scenarios where robots recognize and handle false belief situations in dyadic or
+triadic interactions, and exhibit helping behaviours that account for the
+missing/false beliefs of the human partners.
+
+\cite{devin2016implemented}
+
+\subsubsection{Connections vs. representations}
+\label{connection-representation}
+
+In~\cite{flavell1990developmental}, Flavell relates perspective taking
+\emph{Level 1} to establishing \emph{cognitive connections} (I see, I hear, I
+want, I like, I fear...), in contrast to perspective taking \emph{Level 2} that
+relates to manipulating \emph{representations}.  This is exemplified by
+\emph{appearance-reality} tasks, like the \emph{elephant mask} experiment
+proposed in~\cite{flavell1990developmental}: 3-years old children are not able
+to tell that an experimenter hidden behind a large elephant mask but who speaks
+normally \emph{looks} like an elephant, \emph{sounds} like the experimenter, and
+\emph{really is} the experimenter.  It appears that, while those children are
+able to explicitly manipulate cognitive connections (they know for instance that
+these are largely independent of each other and that they can evolve over time)
+and know as well that their own connections are independent of those of other
+people, they do not think that one concept can \emph{seriously} (\ie non
+playfully) hold several, possibly conflicting, representations.
+
+\subsection{Attention in Robotics}
+
+\section{Methodology: Social Interaction during Free Play}
+
+The study of the emergence of social behaviour is often intrinsically limited by
+the nature of the experimental tasks. We introduce a new
+experimental paradigm based on free play interactions. Where traditional
+socio-cognitive tasks are either toy scenarios (\ie they do not adequately
+mirror real-world situations) or simple, constrained tasks that do not reflect
+the complexity and dynamics of real-world interactions, free play social
+interactions exhibit a rich set of natural cognitive and social dynamics that
+are desirable for this research.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{setup.png}
+    \caption{The free play social interactions sandbox}
+    \label{fig|setup}
+\end{figure}
+
+To mitigate the technical challenges presented by free play interactions (like
+high dimensionality and uncertainties), we have developped a  \emph{free
+play sandbox}~\ref{fig|setup}: open-ended and non-directive activities, yet sufficiently
+defined to be reproducible; focus on abstract socio-cognitive facets (robot
+perception and manipulation are simplified); well suited for qualitative and
+quantitative analysis using metrics like Słowinski’s Individual Motor Signature
+(for behavioural alignment), Ballard’s coding of children’s free-play
+interactions, with-me-ness (for assessment of co-engagment).
+
+\subsection{Free play task}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{sandbox}
+    \caption{Example of a game situation: the child and the robot `build'
+    together a zoo. The animals and the blocks can be dragged over the whole
+    play area, while the background picture is static and purely decorative.}
+    \label{fig|sandbox}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.7\linewidth]{sandbox-release-cheetah}
+    \caption{Possible play situation: the child decides to release the cheetah
+    and creates a story where the cheetah scares away the visitors.}
+    \label{}
+\end{figure}
+
+\subsection{Data Analysis}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{analysis}
+    \caption{Screenshot of the interaction analysis tool: faces of the two
+    children are recorded, as well as any interaction occuring on the
+    interactive table.}
+    \label{fig|analysis}
+\end{figure}
+
+\section{Discussion: Towards a Model of Socio-Cognitive Awareness for Robots}
+
+\subsection{The classical approach}
 
@@
 
-\subsection{From Social Attention to Social Modelling}
+[What is not adequately solved by current techniques?]
 
-The hypothesis that we hereafter develop and turn into a cognitive model is the
-following: mental representations are snapshots of awareness, awareness being
-itself a label for the memory-mediated process of attention.
+[give concrete examples of social situations that are not easily achieveable
+with current (symbolic or not) approaches]
 
-\subsection{Surface Alignment vs Deep Grounding}
-
-Pickering and Garrod~\cite{pickering2006alignment} argue that mutual understanding
-starts mostly with a \emph{superficial alignment} at the level of the linguistic
-representations, due to priming mechanisms, and that this local alignment may --
-in some cases -- lead to a \emph{global alignment} of the semantic level
-(\emph{deep grounding}).  For these authors, the convergence in dialogue, and
-even the repair of some misunderstandings, is explained by this mimetic behavior
-more than by a monitoring of each other's knowledge:
-
-\begin{quote}
-[\ldots] interlocutors do
-not need to monitor and develop full common ground as a regular, constant
-part of routine conversation, as it would be unnecessary and far too costly.
-Establishment of full common ground is, we argue, a specialized and
-non-automatic process that is used primarily in times of difficulty (when
-radical misalignment becomes apparent).~\citep[p.179]{pickering2006alignment}
-\end{quote}
-
-\subsection{Theory of Mind}
-
-We previously covered the state of the research on theory of mind in social
-robotics~\cite{lemaignan2015mutual}.
-
-In~\cite{scassellati2002theory}, Scassellati gave
-an initial account of Leslie's and Baron-Cohen's respective models of the
-emergence of a theory of mind (we discuss them below) from the perspective of
-robotics, but reported implementation work was limited to simple perceptual
-precursors (like face detection or color saliencies detection). Since then,
-research in this field has been focused on applications relying on Flavell's
-\emph{Level 1}~\cite{flavell1977development} perspective-taking, \ie
-perspective-taking that only requires perceptual abilities (``\emph{I see (you do
-not see the book)}''), and actually mostly limited to visual perception (relevant
-work include Breazeal~\cite{breazeal2006using}, Trafton~\cite{Trafton2005} and
-Ros~\cite{Ros2010}).
-
-Based on perspective taking \emph{Level 1} alone, Breazeal
-\etal\cite{breazeal2009embodied} and Warnier \etal\cite{warnier2012when}
-successfully tackled the classical hallmark of theory of mind, the false belief
-experiment (also known as the ``Sally and Anne'' experiment, introduced
-by~\cite{wimmer1983beliefs}, original experimental setting
-by~\cite{baron1985does}). They demonstrated complete human-robot interaction
-scenarios where robots recognize and handle false belief situations in dyadic or
-triadic interactions, and exhibit helping behaviours that account for the
-missing/false beliefs of the human partners.
-
-\cite{devin2016implemented}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Foundations of the Socio-cognitive Awareness for Robots}
-
-\subsection{Connections vs. representations}
-\label{connection-representation}
-
-In~\cite{flavell1990developmental}, Flavell relates perspective taking
-\emph{Level 1} to establishing \emph{cognitive connections} (I see, I hear, I
-want, I like, I fear...), in contrast to perspective taking \emph{Level 2} that
-relates to manipulating \emph{representations}.  This is exemplified by
-\emph{appearance-reality} tasks, like the \emph{elephant mask} experiment
-proposed in~\cite{flavell1990developmental}: 3-years old children are not able
-to tell that an experimenter hidden behind a large elephant mask but who speaks
-normally \emph{looks} like an elephant, \emph{sounds} like the experimenter, and
-\emph{really is} the experimenter.  It appears that, while those children are
-able to explicitly manipulate cognitive connections (they know for instance that
-these are largely independent of each other and that they can evolve over time)
-and know as well that their own connections are independent of those of other
-people, they do not think that one concept can \emph{seriously} (\ie non
-playfully) hold several, possibly conflicting, representations.
+[this examples should be turned into predictions for what our proposal should be
+able to achieve]
 
@@
 
-\subsection{The Attention Schema Theory}
+\subsection{From Social Attention to Social Modelling: the Attention Schema Theory}
+
+The hypothesis that we hereafter develop and turn into a cognitive model is the
+following: \textbf{mental representations are snapshots of awareness, awareness being
+itself a label for the memory-mediated process of attention}.
+
+This extends to social cognition: \textbf{modelling others' mental representations is
+taking snapshots of their current state of awareness}. As we do not have direct
+access to others' process of attention, it has to be mediated. We suggest that
+\textbf{modelling other's state of awareness is mediated by one's own
+attentional system, through joint attention mechanisms}.
 
@@
 awareness'': as such, ``awareness can in principle be verbally reported''.
-
-\subsection{Associative Memory Networks}
-
-In particular, used for behavioural alignment
-
-\subsection{Semantic Networks}
-
-Ontologies and other Knowledge Representation and Reasoning systems.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{The Robotic Social Awareness Model}
-
-\subsection{Overview}
 
@@
 
-\subsection{Attention}
+\paragraph{Attention}
 
@@
 
-\begin{figure}
+\paragraph{Memory-mediated Attention}
+
+\emph{Attention} is modelled as the set of activated units in an associative
+memory network.
+
+Interplay between multi-modal social cues on one hand, and priming of previously activated physical
+entities (objects, agents) and
+
+Attention
+
+\begin{figure}[h!]
     \centering
@@
 
-\subsubsection{Biasing Mechanisms}
+\paragraph{Biasing Mechanisms}
 
@@
 
-\subsection{Social Attention}
+\paragraph{Social Attention}
 
@@
 
-\subsection{What is the contribution, the novelty?}
+\subsection{What Does the Model Predict?}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{What Does the Model Predict?}
+\paragraph{Protodeclarative pointing}
 
-\subsection{Behavioural Alignment}
-\subsection{Verbalization of the (Social) Attentional Process}
-\subsection{False-beliefs}
+Protoimperative vs protodeclarative pointing~\cite{baron1986perceptual}
 
-\subsection{Symbol Grounding}
+\paragraph{Behavioural Alignment}
+\paragraph{Verbalization of the (Social) Attentional Process}
+\paragraph{False-beliefs}
 
-\subsection{Recursive Awareness}
+Picture ordering protocol, with \emph{mechanical}, \emph{behavioural} and
+\emph{intentional} sequences~\cite{baron1986mechanical}
+
+\paragraph{Symbol Grounding}
+
+\paragraph{Recursive Awareness}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{The Functional Pre-Requisites for Social Robots}
+\paragraph{Adapted behaviour to the interacting agent}
 
-\subsection{Social Assessment}
-\subsubsection{Gaze Direction}
-\subsubsection{Facial and Body Expressions}
-\subsubsection{Gestures}
+A robot facing a baby, a 3 years old, a 5 years old, a 13 years old or an adult
+should behave differently.
 
-\subsection{Situation Assessment}
+\paragraph{Parten's stages of play}
 
-Including saliency information
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Evaluation of the Model}
-
-\subsection{Does it perform as well as Symbolic Architectures?}
-\subsection{Comparison with Symbolic Situation Assessment}
+\section{Conclusion}
 

### 2017-04-06 paper.tex
Parten's stages of play

@@
 \subsection{Attention in Robotics}
+
+\subsection{Stages of play}
+
+In developmental psychology, Parten's {\bf stages of
+play}~\cite{parten1932social}:
+
+\begin{enumerate}
+    \item {\bf Solitary (independent) play}: Playing separately from
+        others, with no reference to what others are doing.
+    \item {\bf Onlooker play}: Watching others play. May engage in
+        conversation but not engaged in doing. True focus on the children at
+        play.
+    \item {\bf Parallel play} (adjacent play, social coaction): Playing
+        with similar objects, clearly beside others but not with them (near
+        but not with others.)
+    \item {\bf Associative play}:  Playing with others without
+        organization of play activity. Initiating or responding to
+        interaction with peers. 
+    \item {\bf Cooperative play}: Coordinating one’s behavior with that
+        of a peer. Everyone has a role, with the emergence of a sense of
+        belonging to a group. Beginning of "team work."
+\end{enumerate}
 

### 2017-04-06 paper.tex
Analysis of the freeplay sandbox

@@
 The study of the emergence of social behaviour is often intrinsically limited by
-the nature of the experimental tasks. We introduce a new
-experimental paradigm based on free play interactions. Where traditional
+the nature of the experimental tasks. Where traditional
 socio-cognitive tasks are either toy scenarios (\ie they do not adequately
@@
 interactions exhibit a rich set of natural cognitive and social dynamics that
-are desirable for this research.
+are desirable for this research. We introduce a new
+experimental paradigm based on free play interactions.
 
@@
 
-To mitigate the technical challenges presented by free play interactions (like
-high dimensionality and uncertainties), we have developped a  \emph{free
-play sandbox}~\ref{fig|setup}: open-ended and non-directive activities, yet sufficiently
+To mitigate the technical challenges presented by free play interactions in
+their most unconstrained version (like
+high dimensionality and uncertainties), we have developped a \emph{free play
+sandbox}, Figure~\ref{fig|setup}: a large touchscreen (60cm $\times$ 33cm) is
+used as an interactive surface (\emph{sandtray}
+paradigm~\cite{baxter2012touchscreen}). Two agents (either a pair of children or
+one child and one robot) play together by freely moving on a 2D plane items
+(characters and blocks, see Figure~\ref{fig|sandbox}). A background image
+depicts a generic empty environment, with different symbolic zones (water,
+grass, path, bushes...). The players do not have any particular task to
+complete, they are simply invited to freely play.
+
+children can engage in open-ended and
+non-directive play situations, yet sufficiently
 defined to be reproducible; focus on abstract socio-cognitive facets (robot
@@
 
-\subsection{Free play task}
+\subsection{Analysis of the free play sandbox}
+
+Our free play sandbox elicits a loosely structured play situation. The
+situation is indeed essentially aimless (free play). The goals and play episodes
+that we observe during the interactions are not pre-defined, and essentially
+unknown until they are decided by the players. In this sense, the free play
+sandbox permits and elicits what is fundamentally \emph{unstructured play}.
+
+However, a second look reveals some important structuring elements.
+
+First, the physical bounds of the sandbox (an interactive table) limit the
+play zone to a well defined and relatively small area. As a consequence,
+children are mostly static (they are sitting in front of the table) and their
+primary form of interaction is 2D pick and place of items by drag and dropping
+them on the screen.
+
+Second, the game items themselves structure the game scenarios. Items are either
+iconic characters (animals or children) or plain blocks
+(Figure~\ref{fig|sandbox}). The former have strong semantics associated to them
+(like 'crocodiles like water and eat children'), while the later have little
+semantics associated. The game background, with its recognizable zones, also
+elicit a particular type of games (like building a zoo or pretending we explore
+the savannah).
+
+Finally, the social setting (two players facing each other) implicitely invites
+social interaction: onlooker play, for instance, is unlikely to be observed as
+the children are physically engaged with the game and the other participant
+(they are sitting in front of game, close to each other). The three higher forms
+of social play listed by Parten (parallel play, associative play and cooperative
+play) are however possible with this setup.
+
+Two more elements (the role and place of the experimenter, and the initial
+prompt given to the participant) have an important impact on the shape of the
+play situation. We discuss separately these two points in the description of the
+experimental procedure.
+
+Overall, the free play sandbox supports a \emph{loosely structured} form of play: the
+actual play situations are not known and might change several times during the
+interaction; the game actions, even though all based on a single modality (picking and
+placing game items), are unlimited; the social interactions between participants
+are multi-modal (speech, body postures, gestures, facial expressions, etc.) and
+unconstrainted. However, the broad domain of the play situations and the range of
+possible social interactions is bound by the physical bounds of the play zone
+and the theme of the game items.
+
+\subsection{Forms of play elicited by the sandbox}
+
+\begin{itemize}
+    \item storytelling
+    \item creative play
+    \item pretend play
+\end{itemize}
+
+\paragraph{From unstructured to structured play}
 
@@
 
+\paragraph{Pretence}
+
 \begin{figure}
@@
     \includegraphics[width=0.7\linewidth]{sandbox-release-cheetah}
-    \caption{Possible play situation: the child decides to release the cheetah
-    and creates a story where the cheetah scares away the visitors.}
+    \caption{Possible play situation: the children pretend to release the cheetah
+    and create a story where the cheetah scares away the visitors.}
     \label{}

### 2017-04-06 paper.tex
Desc of the freeplay sandbox + brainstorm on research questions

@@
     \item Can I have real fun playing with my robot?
+    \item A Robot Which is Fun to Play With: Leveraging AI to Participate in Complex
+        Social Situations
     \item A Window Onto Your Mind: Playful Robots Leveraging Attention to
@@
     invited to freely interact with characters and items on an interactive
-    table, telling and acting stories with no explicit preset goal.
+    table, telling and acting stories with no explicit goal set in advance.
 
@@
 
-\section{Methodology: Social Interaction during Free Play}
+\section{Social Interaction during Free Play}
+
+\subsection{A novel methodology}
 
@@
 \begin{itemize}
+    \item dramatic play
     \item storytelling
@@
 
+\subsection{Implementation}
+
+The sandbox is implemented using mainly two frameworks: the Qt's \emph{QtQuick} framework
+for the graphical interface of the game, and the \emph{Robot Operating System}
+(ROS) for the modular implementation of the data processing and behaviour
+generation pipelines.
+
+Figure~\ref{fig|architecture} presents the software architecture of the sandbox.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{freeplay-sandbox-architecture}
+\label{fig|architecture}
+\caption{Architecture of the free play sandbox. Yellow nodes represent
+    hardware (cameras and robot); purple nodes relate to the game interface;
+    red nodes manage the robot; the green node implements the game policy (when
+    the robot is playing); blue nodes relates to visual attention tracking.}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{rviz-sandtray}
+    \caption{The freeplay sandbox, viewed at runtime within ROS RViz
+    visualisation tool.}
+    \label{fig|rviz}
+\end{figure}
+
+\section{Research Questions and Hypotheses}
+
+The mechanisms underlying the social interactions happening during free play
+raise important questions for social robotics.
+
+\subsection{Game Dynamics}
+
+\begin{itemize}
+    \item do we observe a sequence of sub-games?
+    \item how are these sub-games initiated?
+    \item how does A keep track of what B is playing at?
+    \item how do we segment the action flow into meaningful play situations?
+\end{itemize}
+
+\subsection{Social Dynamics}
+
+\begin{itemize}
+    \item does a ``social protocol'' establish?
+    \item if so, what are these social rules?
+    \item implicit vs explicit rule setting?
+    \item what interaction modalities (or combination thereof) are relied upon
+        to establish and maintain this ``social protocol''?
+\end{itemize}
+
+\subsection{Mental representations}
+
+\begin{itemize}
+    \item Does A know what B is thinking of at time t?
+    \item Do the participants perform explicit grounding of their respective mental models
+        (ie, ask the other one what she is doing/thinking/planning to do?)
+    \item How does the implicit grounding take place? attention tracking?
+    \item Do we need at all to know what the other intends to do to actually
+        collaborate? surface alignment vs deep grounding
+    \item Procedural (how do you do it?) vs Semantic
+        grounding (what are you doing? playing at?)
+\end{itemize}
+
+\subsection{Interplay between these questions}
+
+\begin{figure}
+\centering
+
+\resizebox{\textwidth}{!}{%
+\begin{tikzpicture}[
+                font=\sffamily,
+                >=latex,
+                every edge/.style={draw, very thick},
+                node/.style={draw, rounded corners, align=center, inner sep=5pt, fill=black!20},
+                label/.style={midway, align=center, font=\scriptsize\sffamily, fill=white}]
+
+    %%% NODES
+    \node [node] (mentalrepresentations) {Mental Representations};
+    \node [above left=of mentalrepresentations] (gameactions) {game actions};
+    \node [node, above left=of gameactions] (gamedynamics) {Game dynamics};
+    \node [above right=of mentalrepresentations] (socialactions) {social actions};
+    \node [node, above right=of socialactions ] (socialdynamics) {Social Dynamics};
+
+    %%% CONNECTIONS
+    \path (gamedynamics) edge [<->, dashed] node[label, above] {?}(socialdynamics);
+    \path (gamedynamics) edge [<-, bend right] node[label] {mediate}(gameactions);
+    \path (socialactions) edge [->, bend left] node[label] {reflect}(mentalrepresentations);
+    \path (socialactions) edge [->, dashed] node[label,above] {?}(gamedynamics);
+    \path (gameactions) edge [->, dashed] node[label,above] {?}(socialdynamics);
+    \path (gameactions) edge [->, bend right] node[label] {reflect}(mentalrepresentations);
+    \path (socialdynamics) edge [<-, bend left] node[label] {shape}(socialactions);
+
+\end{tikzpicture}
+}
+\label{fig|researchquestions}
+\caption{Game dynamics, social dynamics and mental representations are linked
+    with each other through social actions (implicit and explicit communication)
+    and actions on the environment (mainly game actions)}
+\end{figure}
+
+\subsection{Hypotheses}
+
+\paragraph{Protodeclarative pointing}
+
+Protoimperative vs protodeclarative pointing~\cite{baron1986perceptual}
+
+\paragraph{Behavioural Alignment}
+\paragraph{Verbalization of the (Social) Attentional Process}
+\paragraph{False-beliefs}
+
+Picture ordering protocol, with \emph{mechanical}, \emph{behavioural} and
+\emph{intentional} sequences~\cite{baron1986mechanical}
+
+\paragraph{Symbol Grounding}
+
+\paragraph{Recursive Awareness}
+
+I can describe what I'm aware of, I can also recursively be aware I'm describing
+my state of awareness, and I can verbalize this second order awareness.
+
+This should work \emph{Attention Schema} theory because the attention schema
+is seen as a regular sensory
+representation~\citep[p.55]{graziano2013consciousness}.
+
+\paragraph{Adapted behaviour to the interacting agent}
+
+A robot facing a baby, a 3 years old, a 5 years old, a 13 years old or an adult
+should behave differently.
+
+\paragraph{Parten's stages of play}
+
+\section{Experiments}
+
+\subsection{Procedure}
+
 \subsection{Data Analysis}
@@
 \end{figure}
+
+\subsection{Results}
 
@@
 
-\subsection{What Does the Model Predict?}
-
-\paragraph{Protodeclarative pointing}
-
-Protoimperative vs protodeclarative pointing~\cite{baron1986perceptual}
-
-\paragraph{Behavioural Alignment}
-\paragraph{Verbalization of the (Social) Attentional Process}
-\paragraph{False-beliefs}
-
-Picture ordering protocol, with \emph{mechanical}, \emph{behavioural} and
-\emph{intentional} sequences~\cite{baron1986mechanical}
-
-\paragraph{Symbol Grounding}
-
-\paragraph{Recursive Awareness}
-
-I can describe what I'm aware of, I can also recursively be aware I'm describing
-my state of awareness, and I can verbalize this second order awareness.
-
-This should work \emph{Attention Schema} theory because the attention schema
-is seen as a regular sensory
-representation~\citep[p.55]{graziano2013consciousness}.
-
-\paragraph{Adapted behaviour to the interacting agent}
-
-A robot facing a baby, a 3 years old, a 5 years old, a 13 years old or an adult
-should behave differently.
-
-\paragraph{Parten's stages of play}
-
 \section{Conclusion}

### 2017-04-06 paper.tex
Bits on the socio-cognitive framing of the freeplay sandbox

@@
 non-directive play situations, yet sufficiently
-defined to be reproducible; focus on abstract socio-cognitive facets (robot
+defined to be reproducible;
+
+\subsection{Socio-cognitive Framing}
+
+focus on abstract socio-cognitive facets (robot
 perception and manipulation are simplified); well suited for qualitative and
@@
 interactions, with-me-ness (for assessment of co-engagment).
+
+carefully framed: as far as possible, they ne-
+cessitate only the cognitive capabilities that they ev-
+idence (e.g. if they evidence a purely socio-linguistic
+mechanism, they will not mandate complex action ca-
+pabilities – they may benefit from it, though),
+
+This point is especially important as it allows an
+incremental path: a robot should be able to tackle one or
+several of these task independently of the others, and a re-
+search team may progressively extend their cognitive models
+to incorporate more and more of the cognitive capabilities
+required by the robot to address all of the tasks.
 
@@
 
-\subsection{Hypotheses}
+\subsection{High level Hypotheses}
+
+\begin{itemize}
+    \item autonomous behaviour \emph{ensues from} 
+\end{itemize}
+
+\subsection{Operational Hypotheses}
 

### 2017-04-20 paper.tex
New focus: social dynamics beyond task specifics

@@
 
-\title{The robot who pretend-played with me}
+\title{Generalized Social Dynamics in Human-Robot Interaction}
 
@@
 
+\begin{abstract}
+
+Relying on free play as a prototypical open task, we investigate the social
+dynamics of children dyads playing together. By analyzing them from the
+perspective of artifical socio-cognition and human-robot interaction, we
+evidence key social patterns and show that robots can recognise them,
+interpret them, and (where applicable) generate them in a task-agnostic
+fashion. By subsequently reusing these social patterns in similar
+child-robot free play situations, we show a significant increase of the
+interaction duration.
+
+\end{abstract}
+
+\section{Introduction}
+
 Alternative titles:
@@
 \begin{itemize}
-    \item Can I have real fun playing with my robot?
+    \item The robot who pretend-played with me
+    \item I Guess Your Next Move! Leveraging Social Cues for Robots to
+        Understand Free-play Situations
     \item A Robot Which is Fun to Play With: Leveraging AI to Participate in Complex
@@
         Real-World Social Dynamic
+    \item A Robot Which Adapts its Social Behaviour to its Partner's Age
 
@@
 
-\begin{abstract}
-
-    This article presents how attention tracking enables a robot to first automatically segment
-    a social interaction into meaningful situations (\emph{situation awareness}), and
-    second, to predict the participants' actions.
+This article presents how attention tracking enables a robot to first
+automatically segment a social interaction into meaningful situations
+(\emph{situation awareness}), and second, to predict the participants' actions.
 
@@
 free dramatic play: pairs of children (4-7 years old) or robot and child are
-    invited to freely interact with characters and items on an interactive
-    table, telling and acting stories with no explicit goal set in advance.
+invited to freely interact with characters and items on an interactive table,
+telling and acting stories with no explicit goal set in advance.
 
@@
 behaviours.
-
-\end{abstract}
-
-\section{Introduction}
 
@@
 quantitative analysis using metrics like Słowinski’s Individual Motor Signature
-(for behavioural alignment), Ballard’s coding of children’s free-play
+(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
 interactions, with-me-ness (for assessment of co-engagment).
@@
     \node [node] (mentalrepresentations) {Mental Representations};
-    \node [above left=of mentalrepresentations] (gameactions) {game actions};
-    \node [node, above left=of gameactions] (gamedynamics) {Game dynamics};
-    \node [above right=of mentalrepresentations] (socialactions) {social actions};
-    \node [node, above right=of socialactions ] (socialdynamics) {Social Dynamics};
+    \node [above left=of mentalrepresentations] (gameactions) {\it game actions};
+    \node [node, above=of gameactions] (gamedynamics) {Game dynamics};
+    \node [above right=of mentalrepresentations] (socialactions) {\it social actions};
+    \node [node, above=of socialactions ] (socialdynamics) {Social Dynamics};
 
@@
 
-Protoimperative vs protodeclarative pointing~\cite{baron1986perceptual}
+Protoimperative vs protodeclarative pointing~\cite{baron1989perceptual}
 
